### 2016-11-23 local-capability-machine.tex
updated f1 and f2

\begin{description}
\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
  \begin{itemize}
  \item Set up activation record
  \item Create local enter capability for activation
  \item Clear unused registers
  \item Jump
  \item Upon return: Run activation code
  \end{itemize}
A more detailed description of each of the above steps:
\begin{description}
  \item [Set up activation record]
    \begin{itemize}
    \item Run malloc to get a piece of memory with space for:
    \begin{itemize}
      \item Words in $\bar{r}_{\var{priv}}$
      \item Code return capability (opc)
      \item Activation code
    \end{itemize}
    \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
    \item Adjust the a copy of the current pc to point to the return address and save it to the activation record.
    \item Write the activation code to the activation record.
  \end{itemize}
  \item [Create local enter capability for activation] Make the capability for the activation record point to the beginning of the activation record and restrict it to a local-enter capability. Place this capability in $r_0$.
  \item [Clear unused registers]
    Clear all the register that are not $\pcreg$, $r_0$ or in $\bar{r}_{\var{args}}$.
  \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
  \item [Activation code] The activation code does the following:
    \begin{itemize}
    \item Move the stored ``private'' words in to their respective $\bar{r}_{\var{priv}}$ registers.
%    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
    \item Load the return address to $\pcreg$
    \end{itemize}
\end{description}
\item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.
\item[\texttt{assert$_{\var{flag}}$ $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$. If they are equal, then execution continues. If they are unequal, then the assertion flag named $flag$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
\end{description}
Note:
\begin{itemize}
\item \texttt{call} will fail if we have local capabilities as part of the ``private'' state as it relies on a capability returned by malloc which will not be permit-write-local. Below, we introduce \texttt{scall} which can handle local capabilities in the ``private'' state.
\end{itemize}
\subsection{Stack}
Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma. The stack is accessed through a local $\rwlx$-capability. Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.

@@

   & $\vdots$

@@

\section{Examples}
\label{sec:examples}
\subsection{Simple Examples Using Local Capabilities}
\label{subsec:example-loc-cap}
In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate where a designated assertion flag is set to 1. The assertion flag is assumed to start as 0 and is different from function to function (so in the program below, \texttt{adv} cannot directly use the assertion flag of \texttt{f}). Notice that an assertion will never terminate in a $\failed$ state.
\subsubsection{Simple use of local capabilities}
ML-like program we had before:
\begin{verbatim}
let f = fun adv =>
          let l = 1 in
          adv();
          assert (l == 1)
\end{verbatim}
Assembly program not using stack
\begin{verbatim}
f1: malloc l
    hstore l 1
    call adv()
    assert l 1
    halt
1f:
\end{verbatim}
\begin{lemma}[Correctness lemma for \texttt{f1}]
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}) \\
    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1},\mathtt{1f},\mathtt{f1}) \\
  \end{align*}
if 
  \[
    (\reg_0\update{\pcreg}{c_{\adv}}
          \update{r_1}{c_{f1}}\update{r_2}{c_\malloc},\hs_\adv \uplus \hs_\flag \uplus \hs_{f1} \uplus m_{\malloc} \uplus \hs_{\var{frame}}) \step[n] (\halted,m')
  \]
where 
\begin{align*}
  &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
  &\dom(\hs_{f1}) = [\mathtt{f1},\mathtt{1f}] \\
  &\dom(\hs_\flag) = [\flag,\flag] \\
  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}]} \text{ for all } n
\end{align*}
and $\hs_\flag(\flag) = 0$. $\hs_{f1}$ contains the code of $f1$ including a $\readwrite$-capability for $\hs_\flag$. $\hs_\adv$ is only instructions.
then
\[
  m'(\flag) = 0
\]  
\end{lemma}

Assembly program using the stack:
\begin{verbatim}
f2: push 1
    scall adv()
    pop r1
    assert r1 1
    halt
2f:
\end{verbatim}

>>>>

\begin{description}
\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
  \begin{itemize}
  \item Set up activation record
  \item Create local enter capability for activation
  \item Clear unused registers
  \item Jump
  \item Upon return: Run activation code
  \end{itemize}
A more detailed description of each of the above steps:
\begin{description}
  \item [Set up activation record]
    \begin{itemize}
    \item Run malloc to get a piece of memory with space for:
    \begin{itemize}
      \item Words in $\bar{r}_{\var{priv}}$
      \item Code return capability (opc)
      \item Activation code
    \end{itemize}
    \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
    \item Adjust the a copy of the current pc to point to the return address and save it to the activation record.
    \item Write the activation code to the activation record.
  \end{itemize}
  \item [Create local enter capability for activation] Make the capability for the activation record point to the beginning of the activation record and restrict it to a local-enter capability. Place this capability in $r_0$.
  \item [Clear unused registers]
    Clear all the register that are not $\pcreg$, $r_0$ or in $\bar{r}_{\var{args}}$.
  \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
  \item [Activation code] The activation code does the following:
    \begin{itemize}
    \item Move the stored ``private'' words in to their respective $\bar{r}_{\var{priv}}$ registers.
%    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
    \item Load the return address to $\pcreg$
    \end{itemize}
\end{description}
\item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.
\item[\texttt{assert$_{\var{flag}}$ $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$ (if one of them is an integer, then use that in the comparison). If they are equal, then execution continues. If they are unequal, then the assertion flag named $flag$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
\end{description}
Note:
\begin{itemize}
\item \texttt{call} will fail if we have local capabilities as part of the ``private'' state as it relies on a capability returned by malloc which will not be permit-write-local. Below, we introduce \texttt{scall} which can handle local capabilities in the ``private'' state.
\end{itemize}
\subsection{Stack}
Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma. The stack is accessed through a local $\rwlx$-capability. Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.

@@

@@

\section{Examples}
\label{sec:examples}
\subsection{Simple Examples Using Local Capabilities}
\label{subsec:example-loc-cap}
In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate where a designated assertion flag is set to 1. The assertion flag is assumed to start as 0 and is different from function to function (so in the program below, \texttt{adv} cannot directly use the assertion flag of \texttt{f}). Notice that an assertion will never terminate in a $\failed$ state.
\subsubsection{Simple use of local capabilities}
ML-like program we had before:
\begin{verbatim}
let f = fun adv =>
          let l = 1 in
          adv();
          assert (l == 1)
\end{verbatim}
Assembly program not using stack. Assume that $r_l \not\in \{\pcreg,r_0 \}$ is a register.
\begin{verbatim}
f1: malloc r_l 1
    store r_l 1
    call adv([],[r_l])
    assert r_l 1
    halt
1f:
\end{verbatim}
\begin{lemma}[Correctness lemma for \texttt{f1}]
  For all $n \in \nats$
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\entry,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}) \\
    c_{f1} & \defeq ((\rwx,\glob),\mathtt{f1},\mathtt{1f},\mathtt{f1}) \\
  \end{align*}
if 
  \[
    (\reg_0\update{\pcreg}{c_{\adv}}
          \update{r_1}{c_{f1}}\update{r_2}{c_\malloc},\hs_\adv \uplus \hs_\flag \uplus \hs_{f1} \uplus m_{\malloc} \uplus \hs_{\var{frame}}) \step[n] (\halted,m')
  \]
where 
\begin{align*}
  &\dom(\hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
  &\dom(\hs_{f1}) = [\mathtt{f1},\mathtt{1f}] \\
  &\dom(\hs_\flag) = [\flag,\flag] \\
  &\heapSat[\hs_{\malloc}]{n}{[0 \mapsto \iota_{\malloc}]} \text{ for all } n
\end{align*}
and $\hs_\flag(\flag) = 0$. $\hs_{f1}$ contains the code of $f1$ including a $\readwrite$-capability for $\hs_\flag$. $\hs_\adv$ is only instructions.
then
\[
  m'(\flag) = 0
\]  
\end{lemma}

Assembly program using the stack:
\begin{verbatim}
f2: push 1
    scall adv([],[])
    pop r1
    assert r1 1
    halt
2f:
\end{verbatim}
